---
title-block-banner: images/banner.jpeg
title: "The application of corpus linguistics in analysing the Coronavirus corpus "
author: "Phuong Dang"
format: 
  html:
    toc: true
    toc-title: On this page
    html-math-method: katex
    css: styles.css
    code-fold: true
    theme: journal
    code-link: true
    abstract-title: "Abstract"
number-sections: true
execute: 
  warning: false
abstract: " The goal of this paper is to apply the method of corpus linguistics to the Coronavirus corpus. We will address the research question: how did people change their views about the Corona pandemic from before to after the lockdown. At different time points, our views about a certain event might be different. It is in the best interest of corpus linguistics to explore the impacts of a particular event on society, economy, and culture. The major statistical method conducted in the paper is correspondence analysis (CA). CA is used to analyze the dependence between two (or more) categorical variables. The categorical variables in corpus linguistics are words, text, phrases, etc. CA then will visualize categories of the variables as points in a low-dimension space to interpret the relationship between/among these variables. We, therefore, aim to see if there are existing varieties in the texts and words used before and after the lockdown."
bibliography:
  - citation.bib
  - packages.bib
engire: knitr
prefer-html: true
---

```{r}
#| label: setup
#| include: false
options(digits = 2, knitr.kable.NA = "")
library(tidyverse)
library(kableExtra)
library(here)
library(xml2)
library(glossr)
plot_fonts <- if (!knitr::is_html_output()) "serif" else "sans"
theme_set(theme_minimal(base_size = 16, base_family = plot_fonts))
knitr::write_bib(
  c("base","mclm", "xml2", "ggplot2", "tidyverse"),
  here("packages.bib")
)
```

```{r}
#| label: var
#| include: false
var_percent <- scales::label_percent(scale = 1, accuracy = 0.01)
```

```{css, echo = FALSE}
.justify {
text-align: justify !important
}
```

```{r}
#| include: false
# Load packages
library(tidyverse)
library(mclm)
library(here)
library(kableExtra)
```

## Introduction {#sec_introduction}

::: justify
So many things have changed since the corona pandemic hit in early 2020. Social distancing and travel restrictions are only parts of the global story. Households and consumers are much more cautious about going out than initially thought. Furthermore, organizations around the world have changed their business models overnight, such as companies setting up remote work and universities having to offer online learning. At the beginning of the pandemic, we thought this was just a common disease, so we can rest assured. However, when the number of cases increased day by day and there was no vaccine intervention at that time; Unfortunately, a lot of people have passed away. Then we have to change our perspective on this pandemic.

This paper aims to show how our views have changed before and after the lockdown (the time we were forced to stay at home). A linguistic corpus is designed to reflect the use of language by a linguistic community in a given situation [@inpro]. The paper will start with the introduction of the corpus collection in @sec-collection. We propose the research question in @sec-question. Subsequently, @sec-method presents the literature review of applicable methodology. For example, correspondence analysis (CA) will be the major statistical method, and frequency analysis (FA) and keyword analysis (KA) will be the secondary methods. Next, we apply the FA, KA, and CA respectively, and discuss the results in @sec-discusion to answer the research question. Finally, the paper will end up with a conclusion and further analysis in @sec-conclusion. The programming suite R will be used to perform analysis [@ca_R].

> **COVID-19: Where we've been, where we are, and where we're going:**
>
> A lot can happen in two years. On March 11, 2020, the World Health Organization declared COVID-19 a global pandemic. As the world stares down year three of a crisis that has affected so many lives and livelihoods, we're bringing you a look at how it has affected business and society as well, and what might come next. One of the hardest things to deal with in this type of crisis is being able to go the distance [@quotation].
:::

## Corpus collection {#sec-collection}

::: justify
The selected corpus is [The Coronavirus Corpus.](https://www-english-corpora-org.kuleuven.e-bronnen.be/corona/) The full corpus is currently about 1589 million words in size but in this paper, I will use the [sample_corpus](https://www.corpusdata.org/formats.asp) which consists of around 3.2 million words. The sample corpus includes the corpora which were collected from January to May 2020.

Since April 2020 was a certain time that 3.9 billion people in more than 90 countries and territories have been forced into lockdown due to the Coronavirus pandemic [@entry_lockdown], therefore, I decided to separate my corpus into two subcorpus. The first subfolder contains the corpus of January, February, and March 2020, which represents to the corpus of the before lockdown. The second subfolder contains corpus of April and May 2020 representing the corpus of the after lockdown.
:::

## Research question {#sec-question}

How did our views of the Coronavirus pandemic change before and after the lockdown?

## Methodology {#sec-method}

### Frequency analysis

::: justify
The first corpus technique is frequency analysis. In this approach, we will explore the frequency of words. The subcorpus will be converted to vectors and from there the term frequence of the words or texts can be found [@term]. The most frequently occurring words will be visualized by the bar plots.
:::

### Keyword analysis

::: justify
Secondly, we apply the Keyword Analysis (KA). Keywords are not essentially the words that appear most frequently in the corpus, but they are words identified by statistical comparison between the 'target' corpus the 'reference' corpus. In particular, we want to see if the word appears more often in the *target* corpus as compared to its occurrence in the *reference* corpus. If yes, the word may be a key term of the target corpus. A keyword list consists items (e.g., texts, words) that are either significantly frequent (positive keywords) or infrequent (negative keywords), and is a useful starting point for much linguist analysis [@scott2001comparing].

The keyword analysis is based on the association scores and it regards three criteria:

-   A **frequency** of three or higher in the target corpus.

-   A **PMI score** of two or higher means that the probability of the keyword in the target context is minimum four time higher than its probability in the data taken altogether.

-   A **signed G** score of two or higher such that higher $G^2$ means there is more evidence that two words are dependent but do not distinguish between attraction.

In KA, we choose the file of January of 2020 as the target corpus, and the remaining corpora including files of February, March, April, and May 2020 as the reference corpus. Since January 2020 was when Corona virus appeared, we choose the corpus of January as a target text, and from there we tend to find which keywords are the best to target and provide valuable insights.
:::

### Correspondence analysis

::: justify
Thirdly, we perform CA, which is the major method in this paper. CA is a discovery technique mainly used to explore complex categorical data by identifying patterns of association and disassociation of those data. This technique makes more sense if there are so many differences between configurations [@ca_explore]. By doing CA we investigate how the observed counts in the contingency table deviate from expected counts under statistical independence.

Methodologically, CA produces the contingency table which indicates the difference between the rows and columns of frequencies. Next, it takes the frequency of co-occurring features and then converts them into distances. Particularly, CA computes the eigenvectors of a correlation matrix and generates the Euclidean map. Those vectors will then be plotted into two or three dimensions through a contingency plot (also called biplot). The plot visualizes how things are linked together by distance or proximity.

In more detail in the method of CA, in a contingency table, there are columns and rows containing all the relative frequencies of co-occurrence frequencies that appear simultaneously. Row profiles and column profiles can be created using the same process. For example, row profiles can be calculated by adding up the number of occurrences for each feature in a row, and then dividing each of those occurrences by their sum. The same step applies to column calculation and will give the column profiles. Note that not all co-occurrences are equal in importance, for example, some infrequency features might affect the co-occurrences. Therefore, CA might use the term "mass" (similarly named "weight") to calculate the average score and resolve certain biases [@CA_in_R; @CA_1]. The mathematics to get the row profile and column profile are as follows:

-   Suppose two categorical variables $X$ and $Y$ have an $n_r * n_c$ cross-table with:

    -   $F_{ij}$ the observed count in cell ($i,j$)

    -   $F_{i+}=\sum_{j=1}^{n_c} F_{ij}$ the marginal row count for row $i$

    -   $F_{+j} = \sum_{i=1}^{n_r} F_{ij}$ the marginal column count for column $j$

    -   $n= \sum_{i=1}^{n_r} \sum_{j=1}^{n_c} F_{ij}$ total count

Furthermore, the distance matrix of CA can be generated by conducting the Chi-squared distance measure between two rows $i$ and $i^*$ or between two columns $j$ and $j^*$. The p-value from the Chi-squared test can notice us that CA is or is not statistically significant to fit the data.

-   The strength of the relationship between $X$ and $Y$ can be measured using the Pearson-Chi square statistic: $$ \chi^2 = \sum_{i=1}^{n_r} \sum_{j=1}^{n_c}\frac{(F_{ij} - E_{ij})^2}{E_{ij}} $$ with $E_{ij} = \frac{F_{i+}F_{+j}}{n}$ the expected count if $X$ and $Y$ are statistically independent.

-   If the Pearson-Chi square test indicates that $X$ and $Y$ are statistically dependent, it is relevant to use CA to further study the nature of the association between $X$ and $Y$.

Additionally, inertia is the term used in CA to illustrate the degree of variation. The inertia is calculated based on observed and expected co-occurrence frequencies. High inertia is when column and row profiles have large deviations from their averages. In CA, we investigate the dissimilarity between the profiles, so the higher gain inertia is preferred.

Moreover, in the biplot, two dimensions of the plot represent a certain percentage in the structure of the data variation. These interpreted variance percentages can be seen in the x-axes and y-axes of the biplot.
:::

## Results and Discussions {#sec-discusion}

### Frequency analysis

::: justify
```{r}
#| include: false
# Load packages for frequency analysis
library(udpipe)
library(textrank)
library(tm)
library(SnowballC)
library(wordcloud)
library(RColorBrewer)
```

-   **Most frequency words in March 2020**

In this section, first, we load the corpus of March 2020 by using `readLines(file.choose())`and choose the file of March 2020 `before_lockdown_20-03`. Next, we convert the text to a corpus by `Corpus(VectorSource(text))`, and proceed the corpus in terms of removing numbers, specifying stopwords as a character vector, removing the english common stopwords, and we use `tm_map`function for these steps. The code to get the list of most frequency of words in March 2020 can be found in the *code-fold* below.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| warning: false
# Load file of March 2020
text <- readLines(file.choose())
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| warning: false
# Convert the text as a corpus
docs <- Corpus(VectorSource(text))
toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))
docs <- tm_map(docs, toSpace, "/")
docs <- tm_map(docs, toSpace, "@")
docs <- tm_map(docs, toSpace, "\\|")

# Convert the text to lower case
docs <- tm_map(docs, content_transformer(tolower))
# Remove numbers
docs <- tm_map(docs, removeNumbers)
# Remove english common stopwords
docs <- tm_map(docs, removeWords, stopwords("english"))
# Remove your own stop word
docs <- tm_map(docs, removeWords, c("said", "may","can", "will", "also"))
# specify your stopwords as a character vector
docs <- tm_map(docs, removeWords, c("blabla1", "blabla2")) 
# Remove punctuations
docs <- tm_map(docs, removePunctuation)
# Eliminate extra white spaces
docs <- tm_map(docs, stripWhitespace)

# Convert document to matrix
dtm <- TermDocumentMatrix(docs)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d1 <- data.frame(word = names(v),freq=v)
```

-   **Most frequency words in April 2020**

We load the corpus of April 2020 by `readLines(file.choose())`and choose file `after_lockdown_20-04`. The process to create the list of most frequency of words of April 2020 is similar to the process we use for March 2020.

```{r}
#| include: false
# Load corpus of April 2020
text <- readLines(file.choose())
```

```{r}
#| include: false
#| warning: false

# Load the data as a corpus
docs <- Corpus(VectorSource(text))

toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))
docs <- tm_map(docs, toSpace, "/")
docs <- tm_map(docs, toSpace, "@")
docs <- tm_map(docs, toSpace, "\\|")

# Convert the text to lower case
docs <- tm_map(docs, content_transformer(tolower))
# Remove numbers
docs <- tm_map(docs, removeNumbers)
# Remove english common stopwords
docs <- tm_map(docs, removeWords, stopwords("english"))
# Remove your own stop word
docs <- tm_map(docs, removeWords, c("said", "may","can", "will", "also"))
# specify your stopwords as a character vector
docs <- tm_map(docs, removeWords, c("blabla1", "blabla2")) 
# Remove punctuations
docs <- tm_map(docs, removePunctuation)
# Eliminate extra white spaces
docs <- tm_map(docs, stripWhitespace)
# Text stemming
# docs <- tm_map(docs, stemDocument)

dtm <- TermDocumentMatrix(docs)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d2 <- data.frame(word = names(v),freq=v)

```

In the @tbl-most, we can find the frequency lists of 20 words for March and April 2020. The @fig-bar shows the highest 10 frequent words. However, the most popular words of March and April corpora are not too different. Therefore, in the next section, we will conduct the keyword analysis to see which keywords of the corpus.
:::

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| label: tbl-most
#| layout-ncol: 2
#| tbl-cap: Most frequency words
#| tbl-subcap: 
#|  - "Most frequency words in corpus of March 2020"
#|  - "Most frequency words in corpus of April 2020"

d1 %>%
  head(20) %>%
  as_tibble() %>% 
  kbl(col.names = c("word","freq")) %>%
  kable_minimal(full_width = FALSE) %>% 
  scroll_box(height = "400px")


d2 %>%
  head(20) %>%
  as_tibble() %>% 
  kbl(col.names = c("word","freq")) %>%
  kable_minimal(full_width = FALSE) %>% 
  scroll_box(height = "400px")
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| label: fig-bar
#| layout-ncol: 2
#| fig-cap: Bar plots of 10 most popular words
#| fig-subcap: 
#|   - "Most frequent words in March 2020"
#|   - "Most frequent words in April 2020"

barplot(d1[1:10,]$freq, las = 2, names.arg = d1[1:10,]$word, col ="lightblue", ylab = "Word frequencies")

barplot(d2[1:10,]$freq, las = 2, names.arg = d2[1:10,]$word, col ="lightblue", ylab = "Word frequencies")
```

### Keyword analysis

::: justify
In the keywords analysis, first we load the corpus *corona* by using `{here}` package. We define the target corpus is the corpus of January 2020 `keep_re("before_lockdown_20-01.txt)`, and the reference corpus is the subcorpus of February, March, April, and May 2020. `drop_re("before_lockdown_20-01.txt")`. Next, we build the frequency lists for the target corpus `flist_target` and the reference corpus `flist_ref`.

```{r}
#| include: false
corpus_folder <- here("corona")
fnames_CORONA <- get_fnames(corpus_folder) %>% 
  keep_re("[.]txt")

print(fnames_CORONA, 10, hide_path = corpus_folder)

# store names of target corpus files in fnames_target
fnames_target <- fnames_CORONA %>%
  keep_re("before_lockdown_20-01.txt")%>%
  print(hide_path = corpus_folder)

# store names of reference corpus files in fnames_ref
fnames_ref <- fnames_CORONA %>%
  drop_re("before_lockdown_20-01.txt") %>%
  print(n = 10, hide_path = corpus_folder)

```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| include: false
# build frequency list for target corpus
flist_target <- fnames_target %>%
  freqlist(re_token_splitter = r"--[(?xi)    \s+   ]--", # whitespace as token splitter
           re_drop_token     = r"--[(?xi)  [:\[\]\@\,\.\"\<p>] ]--", # drop tokens with :,[ or ]
           file_encoding     = "UTF-8") %>%
  print()

# build frequency list for reference corpus
flist_ref <- fnames_ref %>%
  freqlist(re_token_splitter = r"--[(?xi)  \s+   ]--",
           re_drop_token     = r"--[(?xi)  [:\[\]\@\,\.\"\<p>]]--",
           file_encoding     = "UTF-8") %>%
  print()
```

```{r}
#| include: false
# calculate scores
scores_kw <- assoc_scores(flist_target, flist_ref)

# print scores, sorted by PMI
print(scores_kw, sort_order = "PMI")

# print scores, sorted by G_signed
print(scores_kw, sort_order = "G_signed")

# print top score satisfy the criteria PMI>2 and G_signed >=2
top_scores_kw <- scores_kw %>% 
  filter(PMI >= 2 & G_signed >= 2)

# print top_scores_kw, sorted by PMI
top_scores_kw %>%
  print(sort_order = "PMI")

# print top_scores_kw, sorted by G_signed
top_scores_kw %>%
  print(sort_order = "G_signed")

# Save the top keywords of target corpus to csv file
top_scores_kw %>%
  write_assoc("January_top_keywords.csv") 
```

Moreover, we create @tbl-pmi to summarize the top 30 keywords of the target corpus regarding to three criteria *Frequency*, *PMI*, and *Signed* $G^2$ , and we sort these values by descending *PMI.* Meanwhile, we make @fig-pmi which shows text close to the position of their datapoints. In this plot, we only plot the datapoints which have the *Signed* $G^2$ larger than 100.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| label: tbl-pmi
#| tbl-cap: Top 30 keywords of target corpus *January 2020* bases on criteria Frequency, PMI, and Signed $G^2$, and sorted by PMI descending

top_scores_kw %>% # also valid for top_scores_colloc
  as_tibble() %>%
  select(type, a, PMI, G_signed) %>% # select 4 columns
  arrange(desc(PMI)) %>%             # sort by PMI (descending) 
  head(30) %>%                       # select top 30 rows
  kbl(col.names = c("Type", "Frequency", "PMI", r"(Signed $G^2$)")) %>% 
  kable_minimal() %>% 
  scroll_box(height = "400px")
```

```{r}
#| include: false
top_scores_df <- as_tibble(top_scores_kw)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| warning: false
#| label: fig-pmi
#| fig-cap: Keyword analysis based on PMI and Signed G

theme_set(theme_minimal(base_size = 15))
g <- top_scores_df %>%
  ggplot(aes(x = PMI, y = G_signed)) +
  labs(x = "PMI", y = "Signed G")

high_G_signed <- top_scores_df %>% 
  filter(G_signed > 100) # extract types with high G_signed

g + geom_point() +
  ggrepel::geom_text_repel(data = high_G_signed, aes(label = type))
```

According to the @tbl-pmi, KA met the criteria of *Frequency, PMI*, and *Signed* $G^2$ . For instance, the keyword *davos* in the target context was five time higher than its probability in all the data taken together. In the @fig-pmi, some keywords are presented such as *davos, wuhan, wef*. The word *davos* refers to the town Davos in Switzerland, where the World Economic Forum takes places every January. The word *Wuhan* which has the *PMI* approximately 3 and *Signed* $G^2$ around 300, indicates the first place covid-19 was spread.
:::

### Correspondence analysis

::: justify
In CA, two main packages are used such as `{mclm}` and `{mc}`. The package `{kableExtra}` is also used to produce the dedicated tables. First, we load the corpus from the corpus folder by using *`here("corona")`*, and then get the *`get-fnames(ct_corpus)`*, and create the short names for the files by *`short_names(fnames)`*. Finally, we generate the vector for each file belonging to subfolder *before lockdown* or *after lockdown*, and also for identifying of categories and colors of biplot.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
# Loading corpus
ct_corpus <- here("corona")
fnames <- get_fnames(ct_corpus)
short_fnames <- short_names(fnames)
sub_corp <- fnames %>%
  re_retrieve_first("/corona/([^/]+)", requested_group = 1)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
# Create short names
set.seed(8541)
tibble(
  fnames = gsub(ct_corpus, "", fnames),
  short_fnames = short_fnames, sub_corp = sub_corp
) %>%
  group_by(sub_corp) %>% 
  kbl() %>% 
  kable_paper()
```

After loading the corpora, we create some functions to perform the CA. The first function is *`compile_frequencies`* such that rows correspond to features, and columns correspond to files. In the *`map`* function, we produce a frequency list for each filename of the features we are interested in. We combine all files in one data frame. Next, we apply *`t()`* to transpose column to row and row to column in the data frame. For cross-tabulations, data is summarized as a certain variable and its levels are relative to another variable (or variables and levels). After performing the transformation, each row now represents to the file, and each column represents to the features (words). Also, the state *`drop_empty_rc()`* is added to check if there are any empty rows and/or empty columns in the data frame, if so, those empty columns and empty rows will be removed.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
# Function complile_frequencies
compile_frequencies <- function(features,
                                fnames,
                                short_fnames,
                                ...) {
  d <- map(setNames(fnames, short_fnames), function(fname) {
    freqlist(fname, ...)[features]
  }) %>%
  bind_cols() %>% 
  data.frame(row.names = features) %>% 
  as.matrix() %>% 
  t() %>% 
  drop_empty_rc()
}
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
# Function get_coords
get_coords <- function(d_ca, sub_corp) {
  texts_df <- row_pcoord(d_ca)[,c(1, 2)] %>% 
    as_tibble(rownames = "text") %>% 
    mutate(Subcorpus = sub_corp)
  
  words_df <- col_pcoord(d_ca)[,c(1, 2)] %>% 
    as_tibble(rownames = "word")
  
  list(
    rows = texts_df,
    cols = words_df
  )
}
```

In the coordinate function, *`get_coords`*, we create the principal coordinates which consist of rows (in files) and columns (in words). By doing it, we will see how similar one file based to other files, how these words are distributed among them, and how the occurrence of words in its file. We will not only check the frequency of words but also use it to create the biplot. Next, we create a data frame that has these words and files. The data frame is in which rows are presented to the files (texts) `texts_df <- row_pcoord(d_ca)`, and columns are presented to features (words) `words_df <- col_pcoord(d_ca)`.

Furthermore, the *`plot_ca`* function is used to make the biplot. The plot shows the dimensions and the position of features (e.g., words, texts) in horizontal and vertical axes. In *ggplot*, the points are represented for *words*, and the colors are represented for *texts*. Also, *`geom_hline`* is horizontal, and *`geom_vline`* is vertical, and by adding *`ìntercept`* $=0$ we will know the center of a plot. The distances between points are relevant to make sure that one unit in *x* axes with one unit in *y* axes, and *`labs`* for the label of axes.
:::

```{r}
#| code-fold: true
#| code-summary: "Show the code"
# Function plot_ca
plot_ca <- function(ca_coords, variances) {
  dim_1 <- sprintf("Dimension 1 (%.2f %%)", variances[[1]])
  dim_2 <- sprintf("Dimension 2 (%.2f %%)", variances[[2]])
  
  ggplot(ca_coords$cols, aes(x = V1, y = V2)) +
    geom_text(aes(label = word), color = "gray60") +
    geom_point(data = ca_coords$rows, aes(color = Subcorpus)) +
    scale_color_manual(values = c("#0000CD","#DC143C")) +
    geom_hline(yintercept = 0, color = "darkgray") +
    geom_vline(xintercept = 0, color = "darkgray") +
    theme_bw(base_size = 12) +
    labs(x = dim_1, y = dim_2) +
    coord_fixed()
}
```

#### Content words from a predefined list as features

::: justify
In this first approach of CA, we use the list of *content words* as usage-features. The difference between categories is defined regarding to the distances between them. When there is more distance between two categories, there are more differences between them.

Initially, we load the `stop_list.txt`which contains content words and then create the frequency list for corpus `freqlist(fnames)`. We keep only the rank of 150 words `keep_bool(ranks(.) <= 150)`.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
# Load content words
stop_list <- read_types(
  file.path("stop_list.txt")
)

features2 <- freqlist(fnames) %>%
  drop_types(stop_list) %>%
  drop_re(" ") %>%
  keep_bool(ranks(.) <= 150) %>%
  as_types()
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| warning: false
# Correspondence analysis
d_content <- compile_frequencies(features2, fnames, short_fnames)
d_ca2 <- ca(d_content)
ca_coords2 <- get_coords(d_ca2, sub_corp)
```

```{r}
#| include: false
d_ca2 <- ca(d_content)
d_ca2
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
chisq.test(d_content)
```

To interpret CA, the first step is to evaluate whether there is a significant dependency between the rows and columns. Specifically, the Pearson Chi-squared test indicates that texts and words are statistically dependent. In other words, the distribution of the words across the categories differs significantly across files. Thus, it is meaningful to use CA to further study the nature of the relation between row and column profiles ($X$ and $Y$). Next, we apply the CA and summarize the values of principal inertias (eigenvalues) in the @tbl-ca. The first two dimension account for $66.35 + 25.01= 91.36\%$ of the total inertia which equals $0.039976 + 0.015068 + 0.003896 + 0.001311 = 0.060251$.

| Principal inertias (eigenvalues): |          |          |          |          |
|-----------------------------------|----------|----------|----------|----------|
| Dimension                         | 1        | 2        | 3        | 4        |
| Value (Inertia)                   | 0.039976 | 0.015068 | 0.003896 | 0.001311 |
| Percentage (%)                    | 66.35    | 25.01    | 6.47     | 2.18     |

: : Correspondence analysis uses the content words {#tbl-ca}

The `summary()` function summarizes the principal row and column coordinates and the following useful statistic for row and column points. The results of `summary()`of the row profiles is shown in @tbl-rowca.

|     | name   | mass | qlt  | inr | k=1  | cor | ctr | k=2  | cor | ctr |
|-----|--------|------|------|-----|------|-----|-----|------|-----|-----|
| 1   | a_2004 | 328  | 801  | 135 | -137 | 751 | 153 | 35   | 50  | 27  |
| 2   | a_2005 | 284  | 825  | 141 | -150 | 751 | 160 | 47   | 74  | 42  |
| 3   | b_2001 | 25   | 542  | 66  | -107 | 73  | 7   | -274 | 470 | 123 |
| 4   | b_2002 | 49   | 977  | 206 | -44  | 8   | 2   | -497 | 970 | 800 |
| 5   | b_2003 | 314  | 1000 | 452 | 294  | 996 | 678 | 19   | 4   | 8   |

: : Summary of CA analysis, with row profiles {#tbl-rowca}

In which:

-   `cor` is squared correlations which indicate the contribution of a dimension to the inertia of a point. The first dimension $(k=1)$ explains the inertia for corpora of March `b_2003`, April `a_2004`, May `a_2005`, while the second dimension $(k=2)$ explains the inertia for corpora of January `b_2001` and February `b_2002`.

-   `qlt` is total quality. According to @tbl-rowca, all total quality values are above 800 in row and also almost high values in columns, which means that for all row and column points inertia is well explained by the two dimensions.
:::

```{r}
#| include: false
# Summary of CA
summary(d_ca2)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| label: fig-content
#| fig-cap: Biplots of CA with content words as features
#| fig-cap-location: bottom

plot_ca(ca_coords2, summary(d_ca2)$scree[,3])
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| label: fig-con-zoom
#| layout-ncol: 2
#| fig-cap: Zoom in biplot of content words
#| fig-subcap: 
#|   - "Zoom in the content words in before lockdown"
#|   - "Zoom in the content words in after lockdown"


### Zoom in the words in before lockdown
plot_ca(ca_coords2, summary(d_ca2)$scree[,3]) +  coord_fixed(xlim = c(0, 0.3), ylim = c(-0.3, 0))


### Zoom in the words after lockdown
plot_ca(ca_coords2, summary(d_ca2)$scree[,3]) +  coord_fixed(xlim = c(-0.5, 0), ylim = c(0, 0.5))
```

::: justify
By using the content words as predefined features, `r var_percent(summary(d_ca2)$scree[1,3])` of variation was captured in the first dimension and `r var_percent(summary(d_ca2)$scree[2,3])` of variation was captured in the second dimension. That means a total of 91.36% of the variation is explained by CA. @fig-content shows there are many varieties in the content of corpus between before and after the lockdown. Furthermore, we do the magnified plots (the zoom-in plot) for each period to see which particular words make these differences.

In the figure of before-lockdown @fig-con-zoom-1, there are words about *confirmed, virus, cases* showing a significant increase in daily corona infections. The words of *contact, medical, public* describe the actions of governments, organizations, and individuals in controlling the pandemic. The word *contact* highly occurred because we already know that the disease spreads through close contact, so the new cases are discovered through investigation of previous contacts. The word *medical* is described our situation then when the vaccine has not been intervened, so we need to act immediately like staying at home and using sanitizer to wash our hands. From there, we can protect ourselves and our communities from the virus. On the other hand, if we do not follow the health advice or rules regarding the corona pandemic, it could lead to more people being hospitalized and our healthcare system could suffer.

In the figure of after-lockdown @fig-con-zoom-2, there is a high occurrence of word *april* indicating a certain period of lockdown has occurred. The key word *lockdown* here means that we should avoid going out or being in close contact with someone. The keyword *worker* indicates that companies need to move from offline to online working.
:::

#### Using bigrams as types

::: justify
Bigrams are combinations of two words, and are combinations of characters. For example, in different tokenization, we do not look at different words, we consider different combinations of characters, but conversely different bigrams. For a specific word as a feature, we can use a specific bigram. For the word occurrences, a few there are frequent, a lot there are infrequent. Also, the longer sequence, the less likely they are to occur. The important step in using bigrams is to convert the **text-based** data frame into a **bigram-based** data frame (*`ngram_size = 2`*). @tbl-bigram shows the highest frequency out of 150 combinations of two words from the bigrams.
:::

```{r}
#| code-fold: true
#| code-summary: "Show the code"
features3 <- fnames %>%
  freqlist(ngram_size = 2) %>%
  drop_re(" ") %>%
  keep_bool(ranks(.) <= 150) %>%
  as_types()
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| label: tbl-bigram
#| tbl-cap: Frequency list of CA using bigram as type
#| tbl-cap-location: bottom
d_bi <- compile_frequencies(
  features3,
  fnames,
  short_fnames,
  ngram_size = 2
  )

kbl(d_bi, booktabs = TRUE) %>% kable_paper()
```

```{r}
#| include: false
# Conduct CA
d_ca_bi <- ca(d_bi)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| label: fig-bigrams
#| fig-cap: Using bigram as type

ca_coords3 <- get_coords(d_ca_bi, sub_corp)
plot_ca(ca_coords3, summary(d_ca_bi)$scree[,3])
```

::: justify
There was `r var_percent(summary(d_ca_bi)$scree[1,3])` variation explained from the first dimension and `r var_percent(summary(d_ca_bi)$scree[2,3])` variation explained from the second dimension. In total, there were 91.45% of variance that CA captures. As we can see in the @fig-bigrams, the distinguish of two subcorpus definitely comes from the second dimension such that the before_lockdown corpus is below the zero-horizontal line, while the after_lockdown corpus is above the zero-horizontal line. We also can see in @tbl-con-1, features below the second dimension (V2) are represented for the corpora of **before-lockdown** such that they have negative value in *y-axes* (-0.35 and -0.24), and features above the second dimension are more characterized for the corpora of **after_lockdown** such that they have positive value in *y-axes* (0.04 and 0.02). Only corpus of March 2020 *before_lockdown_20.03* is separated to other corpora by the first dimension such that its features are located in a positive side of *x-axes* (0.33).
:::

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| label: tbl-con
#| tbl-cap: Subset of texts and content words with principal components.
#| tbl-subcap: 
#|   - "Texts"
#|   - "Features"
#| layout-ncol: 2

set.seed(2022)
ca_coords3$rows %>%  
  slice_sample(n = 5) %>% kbl() %>% 
  kable_paper(full_width = FALSE)

ca_coords3$cols %>%
  slice_sample(n = 10) %>% kbl() %>% 
  kable_paper(full_width = FALSE)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| label: fig-bigrams-zoom
#| layout-ncol: 2
#| fig-cap: Zoom in biplot of bigrams
#| fig-subcap: 
#|   - "Zoom in biplot of before lockdown"
#|   - "Zoom in biplot of after lockdown"


### Zoom out the words in before lockdown
plot_ca(ca_coords3, summary(d_ca_bi)$scree[,3]) +
  coord_fixed(xlim = c(-0.2,0.5), ylim = c(-0.4, -0.1))

### Zoom out the words after lockdown
plot_ca(ca_coords3, summary(d_ca_bi)$scree[,3]) +
  coord_fixed(xlim = c(-0.3,0.2), ylim = c(0, 0.3))
```

::: justify
In the one hand, the figure of *before-lockdown* @fig-bigrams-zoom-1 shows the combination words of *cases-of, the-new, the-virus* which describes the pandemic situation at that time such as having a large number of new cases each day. The figure of *after-lockdown* @fig-bigrams-zoom-2, on the other hand, shows a combination of words of *social-distancing, at-home* representing the solutions we used at that time to prevent exposure to the virus.

> Social distancing is the term used to describe a series of actions or measures designed to limit or control the infection in order to stop or slow down the spread of a contagious disease. The aim is to reduce the possibility of contact between people who have been infected and others, and thus minimize the spread of the disease [@social_distancing].
:::

```{r}
#| code-fold: true
#| code-summary: "Show the code"

# Zoom in the bottom right of biplot (words from before lockdown)
bottom_right <- ca_coords3$cols %>%
  filter(V1 >= 0 & V2 <= 0) %>% 
  mutate(dist_to_center = sqrt(V1^2 + V2^2)) %>% 
  arrange(desc(dist_to_center)) %>% 
  pull(word)

bottom_left <- ca_coords3$cols %>%
  filter(V2 >= 0 & V1 <= 0) %>% 
  mutate(dist_to_center = sqrt(V1^2 + V2^2)) %>% 
  arrange(desc(dist_to_center)) %>% 
  pull(word)
```

::: justify
Furthermore, we collect the words showing in the zoom-in plots. For example, there are **`r length(bottom_right)`** words in @fig-bigrams-zoom-1 , and the thirteen words which might belong to the ***before lockdown*** corpora are `r glue::glue_collapse(paste0("*", head(bottom_right, 13), "*"), sep = ", ", last = " and ")`. In the other side, there are **`r length(bottom_left)`** words in @fig-bigrams-zoom-2 , and the thirteen words which seem to belong to the ***after lockdown*** corpora are `r glue::glue_collapse(paste0("*", head(bottom_left, 13), "*"), sep = ", ", last = " and ")`.
:::

```{r}
#| include: false
d_ins_bi <- head(ca_coords3,20)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| label: tbl-bi-ins
#| layout-ncol: 2
#| tbl-cap: Inspect extreme features of bigrams
#| tbl-subcap: 
#|  - "Inspect extremes by V2 in ascending "
#|  - "Inspect extremes by V2 in descending "


d_ins_bi$cols %>% 
  as_tibble() %>% 
  arrange(V2) %>%
  kbl(col.names = c("word","V1", "V2")) %>%
  kable_minimal(full_width = FALSE) %>% 
  scroll_box(height = "400px")


d_ins_bi$cols %>% 
  as_tibble() %>% 
  arrange(desc(V2)) %>%
  kbl(col.names = c("word","V1", "V2")) %>%
  kable_minimal(full_width = FALSE) %>% 
  scroll_box(height = "400px")

```

```{r}
#| include: false
before <- ca_coords3$cols %>%
  arrange(V2) %>%
  head(20) %>% pull(word) 

after <- ca_coords3$cols %>%
  arrange(desc(V2)) %>%
  head(20) %>% pull(word) 

```

::: justify
Because two subcorpus are separated by the second dimensions so first we do `arrange(V2)` to find the words belonging to the corpus *before lockdown*, and second we do `arrange(desc(V2))` to get the words belonging to the corpus *after lockdown*. The most characteristic features of **before lockdown** are `r glue::glue_collapse(paste0("*", before, "*"), sep = ", ", last = " and ")`, while those most characteristic features of **after lockdown** are `r glue::glue_collapse(paste0("*", after, "*"), sep = ", ", last = " and ")`.
:::

#### Using trigrams as types

::: justify
In this section, we use trigrams as types for conducting of CA. Trigrams is a combination in sequence of three words, and combinations of characters. The longer sequence (e.g., sequence of three words here), the less frequently they appear. Also, to proceed the trigrams, we need to convert the **text-based** data frame into a **trigram-based** data frame (*`ngram_size = 3`*).
:::

```{r}
#| code-fold: true
#| code-summary: "Show the code"
features4 <- fnames %>%
  freqlist(ngram_size = 3) %>%
  drop_re(" ") %>%
  keep_bool(ranks(.) <= 150) %>%
  as_types()

d_tri <- compile_frequencies(
  features4,
  fnames,
  short_fnames,
  ngram_size = 3
  )

```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| label: tbl-tri
#| tbl-cap: Frequency list of trigrams
#| tbl-cap-location: bottom
kbl(d_tri, booktabs = TRUE) %>% kable_paper()
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| label: fig-tri
#| fig-cap: Biplot of trigram
#| fig-cap-location: bottom
d_ca_tri <- ca(d_tri)
ca_coords4 <- get_coords(d_ca_tri, sub_corp)
plot_ca(ca_coords4, summary(d_ca_tri)$scree[,3])
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| label: fig-tri-zoom
#| layout-ncol: 2
#| fig-cap: Zoom in biplot of trigrams
#| fig-subcap: 
#|   - "Zoom in the features in before lockdown of trigrams"
#|   - "Zoom in the features in after lockdown of trigrams"

### Zoom out the words in before lockdown
plot_ca(ca_coords4, summary(d_ca_tri)$scree[,3]) +
  coord_fixed(xlim = c(-0.5,0), ylim = c(-0.8, -0.3))

### Zoom out the words after lockdown
plot_ca(ca_coords4, summary(d_ca_tri)$scree[,3]) +
  coord_fixed(xlim = c(-0.5,0), ylim = c(0, 0.4))
```

```{r}
#| include: false
d_ins_tri <- head(ca_coords4,20)

```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| label: tbl-tri-inspect
#| layout-ncol: 2
#| tbl-cap: Inspect extremes by V2
#| tbl-subcap: 
#|  - "Inspect extremes by ascending of V2"
#|  - "Inspect extremes by descending of V2"

d_ins_tri$cols %>% 
  as_tibble() %>% 
  arrange(V2) %>%
  kbl(col.names = c("word","V1", "V2")) %>%
  kable_minimal(full_width = FALSE) %>% 
  scroll_box(height = "400px")


d_ins_tri$cols %>% 
  as_tibble() %>% 
  arrange(desc(V2)) %>%
  kbl(col.names = c("word","V1", "V2")) %>%
  kable_minimal(full_width = FALSE) %>% 
  scroll_box(height = "400px")

```

::: justify
@fig-tri shows that the subcorpus of *before lockdown* and *after lockdown* are seemly distinguished by the second dimension (V2), so we do inspection of extreme features based on V2. In @tbl-tri-inspect-1, we sort the combination of words in three by the ascending of second dimension (V2), while in @tbl-tri-inspect-2 is descending of V2. As we can see in the @fig-tri-zoom-2, some words in sequence are *the_covid-19_pandemic, the_univeristy_of, the_coronavirus_outbreak, in_response_to*. The first dimension accounts for are `r var_percent(summary(d_ca_tri)$scree[1,3])` and the second dimension accounts for `r var_percent(summary(d_ca_tri)$scree[2,3])` of the total inertia.
:::

### Remarks

::: justify
In the first statistical approach, frequency analysis, we can see the highest occurrence frequency of 20 words in the corpus for March and April. Some of the most popular words are *coronavirus, covid, now, time, public, government, people, health, new.* Because these words mainly reflect the context of the *Coronavirus*, so it is clear that it has a high probability of appearing in the corpus. In the analysis of keywords, we do not focus on the frequency of occurrence of words, but often tend to compare the *target* corpus and the *reference* corpus. The result of KA all satisfies the three criteria of *Frequency, PMI,* and *Signed* $G^2$.

In CA, the first step is to evaluate whether there is a significant dependency between the rows and columns. We used the `chi-square statisitc` for examining the association between row and column variables. Since p-value of *Pearson Chi-squared test* is statistically significant, so we rejected the null hypothesis $(H_0)$ and concluded that applying CA is meaningful to investigate the association between row and column profiles. Moreover, in CA, we applied *n-gram* techniques. First n-gram is ***Unigram*** in which we used the *content words* as features for CA. Based on biplot, we could see special words such as *break, covid-19, pandemic, lockdown*. Next, we made the second *n-gram* as ***Bigram***. In this *n-gram*, the text is the combinations of two words, so the size of frequency list can be less than a unigram (e.g., only one word). Finally, we proceed to another *n-gram* called ***Trigram***. In this *n-gram*, the text (item) is the combinations of three words, and obviously, the size of frequency list of word combinations is smaller than *bi-grams* or *unigram*. Regarding the *n-gram* approaches in CA, more than $90\%$ of the variation is explained by the first two dimensions. Based on the results and discussions of these methods, we illustrate that there are dissimilarities in the words and texts used in the corpora of before and after lockdown because two subcorpus were separated by first two dimensions. We, therefore, can state that the views of people about the covid pandemic have changed when lockdown happened.
:::

::: column-margin
The null hypothesis of Pearson Chi-squared test is: $H_0$ : $X$ and $Y$ are statistically independent.

The alternative hypothesis is:

$H_a$: $X$ and $Y$ are statistically dependent.
:::

## Conclusion {#sec-conclusion}

::: justify
In this paper, we explore the change in people's viewpoint about the corona pandemic. Through various statistical techniques, we generate the frequency lists of words, the keywords, and the correspondence analysis for investigating varieties. Based on the result of our analysis, we conclude that human views have changed since the coronavirus lockdown. However, in order to firmly emphasize the impacts of coronavirus in our lives from the beginning of the pandemic to the present, further investigations such as using a larger corpus, or taking other approaches may be necessary like collocation, logistic regression, and clustering analysis.
:::

## Acknowledgment {.unnumbered .unlisted .appendix}

::: justify
I am thankful for the teaching and the dedicated feedback offered by teacher **Mariana Montes**. Learning from my mistakes and errors help me not only gain knowledge in my study but also improve myself in the further.
:::
